---
title: "🛍️ Fashion Retail Analysis"
author: "Yvonne Kirschler"
subtitle: "NDK HF Transfer Project · Created with R and Quarto"
format:
  html:
    toc: true
    toc-depth: 2
    toc-title: "Navigation"
    toc-location: left
    toc-floating: true
    number-sections: true
    code-fold: true
    fig-align: center
    theme: cosmo
    df-print: paged
    highlight-style: github
    embed-resources: true
mainfont: "Helvetica Neue"
fontsize: 1.05em
editor: visual
---

```{r setup, include=TRUE, collapse=TRUE, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(gt)
library(broom)
library(glue)
library(skimr)
library(scales)
```

*All analyses were conducted using R (≥ 4.4) and rendered with Quarto.\
All code is visible and reproducible in accordance with NDK HF guidelines.*

> *This report was written independently. ChatGPT (OpenAI) was used to support structure, phrasing, and quality control.\
> All analysis, implementation and interpretation were carried out entirely by the author.*

## Research Question

**Which factors influence revenue per transaction in fashion retail?**

This project investigates:

-   Whether discounts reduce average transaction revenue\
-   How product categories affect sales\
-   Whether payment method has a measurable impact\
-   How these effects can be validated statistically

## Background

This report is based on a synthetic dataset that closely simulates global fashion retail operations.\
The goal is to analyze factors that impact revenue per transaction – such as discount usage, product type, and payment method.\
Statistical methods like confidence intervals and regression models are applied to uncover these patterns.

## Dataset Overview

The dataset includes six structured CSV files:

-   `transactions.csv`: Transaction records (\~6.4 million rows)\
-   `customers.csv`: Customer demographic information\
-   `products.csv`: Product categories and descriptions\
-   `stores.csv`: Store locations and metadata\
-   `employees.csv`: Staff assignments\
-   `discounts.csv`: Promotional campaigns

⚠️ The data is stored locally and is not included in the repository.

## Data Preparation

```{r load-data}
transactions <- read_csv("data/transactions.csv", show_col_types = FALSE) |> clean_names()
products     <- read_csv("data/products.csv", show_col_types = FALSE)     |> clean_names()
discounts    <- read_csv("data/discounts.csv", show_col_types = FALSE)    |> clean_names()
stores       <- read_csv("data/stores.csv", show_col_types = FALSE)       |> clean_names()
customers    <- read_csv("data/customers.csv", show_col_types = FALSE)    |> clean_names()
employees    <- read_csv("data/employees.csv", show_col_types = FALSE)    |> clean_names()

transactions <- transactions |> mutate(discount_applied = if_else(discount > 0, "Yes", "No"))
products     <- products |> mutate(category = as_factor(category))
```

### Missing Values

We check for missing values across the main datasets.\
This is important to ensure valid calculations and prevent bias in averages or model outputs.

```{r missing-values}
sapply(list(transactions = transactions, products = products, discounts = discounts), \(df) sum(is.na(df)))
```

## Exploratory Analysis

### Top 10 Stores by Revenue

```{r}
revenue_by_store <- transactions |> 
  group_by(store_id) |> 
  summarise(total_revenue = sum(line_total, na.rm = TRUE)) |> 
  arrange(desc(total_revenue)) |> 
  slice_head(n = 10) |> 
  mutate(store_label = paste("Store", store_id))

ggplot(revenue_by_store, aes(x = fct_reorder(store_label, total_revenue), y = total_revenue)) +
  geom_col(fill = "#6BCABA") +
  coord_flip() +
  scale_y_continuous(labels = label_comma()) +
  labs(title = "Top 10 Stores by Total Revenue", x = "Store ID", y = "Total Revenue")
```

### Monthly Revenue Trend

```{r}
transactions_by_month <- transactions |> 
  mutate(month = floor_date(date, unit = "month")) |> 
  group_by(month) |> 
  summarise(total_revenue = sum(line_total, na.rm = TRUE))

ggplot(transactions_by_month, aes(x = month, y = total_revenue)) +
  geom_line(color = "#4FB8A8", linewidth = 1.2) +
  scale_y_continuous(labels = label_comma()) +
  labs(title = "Monthly Revenue Trend", x = "Month", y = "Revenue")
```

### Average Revenue by Product Category

```{r}
revenue_by_category <- transactions |> 
  left_join(products, by = "product_id") |> 
  group_by(category) |> 
  summarise(avg_revenue = mean(line_total, na.rm = TRUE)) |> 
  arrange(desc(avg_revenue))

ggplot(revenue_by_category, aes(x = fct_reorder(category, avg_revenue), y = avg_revenue)) +
  geom_col(fill = "#7ACCB8") +
  coord_flip() +
  scale_y_continuous(labels = label_comma()) +
  labs(title = "Average Revenue by Product Category", x = "Product Category", y = "Average Revenue")
```

### Revenue Distribution: Discount vs. No Discount

```{r}
transactions |> 
  filter(line_total > 0) |> 
  ggplot(aes(x = discount_applied, y = line_total)) +
  geom_boxplot(fill = "#CBD5D1", color = "#4FB8A8") +
  scale_y_log10(labels = label_comma()) +
  labs(title = "Revenue Distribution by Discount Applied", x = "Discount Applied", y = "Revenue (log10)")
```

### Revenue vs. Store Size (Scatterplot)

```{r}
revenue_by_store_size <- transactions |> 
  group_by(store_id) |> 
  summarise(total_revenue = sum(line_total, na.rm = TRUE)) |> 
  mutate(store_id = as.numeric(as.character(store_id))) |> 
  left_join(stores, by = "store_id")

ggplot(revenue_by_store_size, aes(x = number_of_employees, y = total_revenue)) +
  geom_point(alpha = 0.6, color = "#6BCABA", size = 2) +
  scale_y_continuous(labels = label_comma()) +
  labs(title = "Revenue by Store Size", x = "Number of Employees", y = "Total Revenue")
```

## Confidence Interval using `qnorm()`

We calculate 95% confidence intervals using the standard formula:

``` r
qnorm(c(0.025, 0.975), mean, sd / sqrt(n))
```

```{r ci-table}
ci_discount <- transactions |> 
  group_by(discount_applied) |> 
  summarise(
    n = n(),
    mean = mean(line_total, na.rm = TRUE),
    sd = sd(line_total, na.rm = TRUE),
    lower_ci = qnorm(0.025, mean, sd / sqrt(n)),
    upper_ci = qnorm(0.975, mean, sd / sqrt(n))
  )

ci_discount |> 
  rename(std_dev = sd) |> 
  gt() |> 
  tab_header(title = "95% Confidence Intervals for Revenue per Transaction") |> 
  fmt_number(columns = 2:6, decimals = 2)
```

## Linear Regression Model

```{r regression}
model_data <- transactions |> 
  filter(line_total > 0) |> 
  left_join(products, by = "product_id") |> 
  mutate(
    discount_applied = as_factor(discount_applied),
    category         = as_factor(category),
    payment_method   = as_factor(payment_method)
  ) |> 
  select(line_total, discount_applied, category, payment_method)

model_formula <- line_total ~ discount_applied + category + payment_method
model_lm <- lm(model_formula, data = model_data)
tidy_model <- tidy(model_lm)

summary(model_lm)
```

```{r regression-table}
tidy_model |> 
  arrange(desc(abs(estimate))) |> 
  gt() |> 
  tab_header(title = "Regression Results: Line Revenue Model") |> 
  fmt_number(columns = 2:5, decimals = 4)
```

## Summary Table by Payment Method

```{r payment-summary}
transactions |> 
  filter(line_total > 0) |> 
  group_by(payment_method) |> 
  summarise(
    avg_revenue = mean(line_total, na.rm = TRUE),
    n = n()
  ) |> 
  arrange(desc(avg_revenue)) |> 
  gt() |> 
  tab_header(title = "Average Revenue per Transaction by Payment Method") |> 
  fmt_number(columns = 2, decimals = 2)
```

## Conclusion

This analysis shows that revenue per transaction in fashion retail is not random, but driven by several statistically measurable factors:

-   **Discounts** significantly lower the revenue per transaction (approx. −64 CHF)\
-   **Product categories** like “Masculine” result in higher average sales than others\
-   **Payment methods** like credit cards produce slightly higher revenues

The linear model explains about **2.7% of total variance** – a realistic value for transactional business data.

### Practical Implications:

-   Use discounts strategically to avoid unnecessary revenue loss\
-   Optimize the product mix based on data-driven category insights\
-   Consider payment methods when planning pricing or promotions

> *This report fulfills all requirements of the NDK HF Transfer Project and is fully reproducible.*
